Covariates analysis of MSBB RNASeq data of region BM 36 (realigned from synapse BAM with Picard-STAR-HTSeq)

### 2016-10-25 code/Rmd/MSSM_BM_36_RNASeq_Reprocessed_SynBAM_StarHTSeq.Rmd
Added markdown to process BM36 from synapse BAM with picard star htseq pipeline

---
title: "Covariates analysis of MSBB RNASeq data of region BM 36 (realigned from FASTQ extracted source BAM in minerva with STAR-HTSeq)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knitr)
library(githubr)

@@

knit2synapse::knitToFolderEntity(file = "./MSSM_BM_36_RNASeq_Realigned_StarHTSeq2.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB BM 36 RNASeq RealignedFromSource BAM Picard StarHTSeq')
```

@@

thisFileName <- 'MSSM_BM_36_RNASeq_Realigned_StarHTSeq2.Rmd'

@@

Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_ID <- 'syn7254146'
ALL_USED_IDs <- COUNT_ID
COUNT <- read.table(synGet(COUNT_ID)@filePath, quote = '', header = T, sep = '\t')

# Get sample ids
SampleID = data.frame(SampleID = colnames(COUNT),
                      ID = colnames(COUNT)) %>% 
  tidyr::separate(ID, c('A','B','ID', 'C'), sep = '_')

SampleID = SampleID %>%
  dplyr::select(SampleID, ID) %>%
  dplyr::mutate(ID = as.numeric(ID),
                Sample.ID = SampleID,
                SampleID = paste0('BM_36_',ID))

# Sum counts from multiple lanes
COUNT = dlply(SampleID, .(ID), .fun = function(x, count){
  rowSums(count[,x$Sample.ID,drop=F]) %>%
    rownameToFirstColumn('Gene.ID') %>%
    plyr::rename(c('DF' = paste0('BM_36_',unique(x$ID)))) %>%
    dplyr::select(-Gene.ID)
}, COUNT) %>%
  Reduce(cbind, .)
COUNT = data.matrix(COUNT)

@@

# Get technical metadata
METADATA_TECHNICAL_ID <- 'syn5898489'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TECHNICAL_ID
METADATA_TECHNICAL <- read.table(synGet(METADATA_ID)@filePath,sep='\t',header=T, row.names=1) %>%
  rownameToFirstColumn('OldBarcode')

@@

METADATA = left_join(METADATA_TECHNICAL %>%
                       dplyr::mutate(ID = gsub('hB_RNA_','',NewBarcode),
                                     ID = as.numeric(ID)),
                     METADATA_CLINICAL %>%
                       dplyr::select(BB, BM36.Fnl, PMI, RACE, 
                                     AOD, CDR, SEX, NP.1, PlaqueMean, bbscore) %>%
                       dplyr::rename(BM_36 =  BM36.Fnl) %>%
                       tidyr::gather(BrainRegion, ID, -BB, -PMI, -RACE, 
                                     -AOD, -CDR, -SEX, -NP.1, -PlaqueMean, -bbscore) %>%
                       filter(!is.na(ID)) %>%
                       dplyr::mutate(ID = as.numeric(ID))) %>%
  inner_join(SampleID %>% dplyr::select(-Sample.ID)) %>%
  filter(!is.na(SampleID)) %>%
  filter(!(ID %in% c(335, 338, 339, 341))) %>%
  unique
rownames(METADATA) = METADATA$SampleID

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "MSBB BM 36 RNASeq RealignedFromSource BAM Picard StarHTSeq", parentId = parentId)
CODE <- synStore(CODE)

@@

# Store logcpm values
write.table(VOOM.ADJUSTED.GENE_EXPRESSION$E, file = 'logCPM.tsv', sep = '\t', row.names=T, quote=F)
exp1_obj = File('logCPM.tsv', name = 'Raw logCPM', parentId = CODE$properties$id)
exp1_obj = synStore(exp1_obj, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

```
|  *Results*                |  *SynapseID*                    |
|  -------                  |   ---------                     |
|  Covariates               |  `r covar_obj$properties$id`    |
|  logCPM                   |  `r exp1_obj$properties$id`     |
|  Residual Expression      |  `r expr_obj$properties$id`     |
|  Differential Expression  |  `r dexp_obj$properties$id`     |

>>>>

---
title: "Covariates analysis of MSBB RNASeq data of region BM 36 (realigned from synapse BAM with Picard-STAR-HTSeq)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knitr)
library(githubr)

@@

knit2synapse::knitToFolderEntity(file = "./MSSM_BM_36_RNASeq_Reprocessed_SynBAM_StarHTSeq.Rmd",
                                 parentId ="syn5570248",
                                 entityName = "MSBB BM 36 RNASeq RealignedFromSynBAM PicardStarHTSeq")
```

@@

thisFileName <- 'MSSM_BM_36_RNASeq_Reprocessed_SynBAM_StarHTSeq.Rmd'

@@

Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_ID <- 'syn7425130'
ALL_USED_IDs <- COUNT_ID
COUNT <- read.table(synGet(COUNT_ID)@filePath, quote = '', header = T, sep = '\t')

# Get sample ids
ind.reseq = grep('resequenced', colnames(COUNT))
ind.all = setdiff(1:dim(COUNT)[2], ind.reseq)

SampleID = rbind(data.frame(SampleID = colnames(COUNT)[ind.all],
                      ID = colnames(COUNT)[ind.all]) %>% 
  tidyr::separate(ID, c('A','B','ID', 'C'), sep = '_') %>%
  dplyr::mutate(ID = gsub('.accepted','',ID)) %>%
  dplyr::select(SampleID, ID) %>%
  dplyr::mutate(ID = as.numeric(ID),
                Sample.ID = SampleID,
                SampleID = paste0('BM_36_',ID)),
  data.frame(SampleID = colnames(COUNT)[ind.reseq],
             ID = colnames(COUNT)[ind.reseq]) %>% 
  tidyr::separate(ID, c('A','B','ID', 'C'), sep = '_') %>%
  dplyr::mutate(ID = gsub('.accepted','',ID)) %>%
  dplyr::select(SampleID, ID) %>%
  dplyr::mutate(ID = as.numeric(ID),
                Sample.ID = SampleID,
                SampleID = paste0('BM_36_Reseq_',ID)))

# Sum counts from multiple lanes
# COUNT = dlply(SampleID, .(ID), .fun = function(x, count){
#   rowSums(count[,x$Sample.ID,drop=F]) %>%
#     rownameToFirstColumn('Gene.ID') %>%
#     plyr::rename(c('DF' = paste0('BM_36_',unique(x$ID)))) %>%
#     dplyr::select(-Gene.ID)
# }, COUNT) %>%
#   Reduce(cbind, .)
COUNT = data.matrix(COUNT)

rownames(SampleID) = SampleID$Sample.ID
colnames(COUNT) = SampleID[colnames(COUNT), 'SampleID']

@@

# Get technical metadata
METADATA_TECHNICAL_ID <- 'syn5898489'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TECHNICAL_ID
METADATA_TECHNICAL <- read.table(synGet(METADATA_TECHNICAL_ID)@filePath,sep='\t',header=T, row.names=1) %>%
  rownameToFirstColumn('OldBarcode')

@@

METADATA = left_join(METADATA_TECHNICAL %>%
                       dplyr::mutate(ID = gsub('hB_RNA_','',NewBarcode),
                                     ID = as.numeric(ID)),
                     METADATA_CLINICAL %>%
                       dplyr::select(BB, BM36.Fnl, PMI, RACE, 
                                     AOD, CDR, SEX, NP.1, PlaqueMean, bbscore) %>%
                       dplyr::rename(BM_36 =  BM36.Fnl) %>%
                       tidyr::gather(BrainRegion, ID, -BB, -PMI, -RACE, 
                                     -AOD, -CDR, -SEX, -NP.1, -PlaqueMean, -bbscore) %>%
                       filter(!is.na(ID)) %>%
                       dplyr::mutate(ID = as.numeric(ID))) %>%
  inner_join(SampleID %>% dplyr::select(-Sample.ID)) %>%
  filter(!is.na(SampleID)) %>%
  unique
rownames(METADATA) = METADATA$SampleID

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "MSBB BM 36 RNASeq RealignedFromSynBAM PicardStarHTSeq", parentId = parentId)
CODE <- synStore(CODE)

@@

# Store logcpm values
write.table(VOOM.ADJUSTED.GENE_EXPRESSION$E, file = 'logCPM.tsv', sep = '\t', row.names=T, quote=F)
exp1_obj = File('logCPM.tsv', name = 'Raw logCPM', parentId = CODE$properties$id)
exp1_obj = synStore(exp1_obj, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
```
|  *Results*                |  *SynapseID*                    |
|  -------                  |   ---------                     |
|  Covariates               |  `r covar_obj$properties$id`    |
|  Residual Expression      |  `r expr_obj$properties$id`     |
|  Differential Expression  |  `r dexp_obj$properties$id`     |
